---
title: ""
format: dashboard
---

# Syllabus and Course Links

```{r include=F}
#| label: setup

options(scipen=100)                         # suppress scientific notation
                                            # install pacman if needed
if (!require("pacman")) install.packages("pacman", repos = "http://lib.stat.cmu.edu/R/CRAN/")
pacman::p_load(pacman, tidyverse, ggthemes, gridExtra, magrittr,
               kableExtra, gt)

#pacman::p_loaded()

```

```{r echo=F}
#| label: created formatted date
# create variables for inline code
today <- Sys.Date()                                                # saves today's date
today_fmt <- stamp("Saturday, January 1, 1999", quiet = T)(today)  # formats date

```

## Row {height="30%"}

**Professor Pooler (pspooler\@syr.edu) - Office: WSOM Room
518**

- In-Person Hours: Tue. Thu. 3:40 PM - 4:45 PM
- Zoom Hours: Mon. Wed. 4:30 PM - 5:30 PM - [**Zoom Hrs.
  Link**](https://syracuseuniversity.zoom.us/j/96889308509){target="_blank"}
- Also available by appointment on Zoom or in-person

## Row {height="70%"}

[**Fall 2026
Syllabus**](https://docs.google.com/document/d/1ycqRSNZWkvodKbACx_fq_Bx8cS-HBdHsLDcDNTIoz6I/edit?usp=sharing){target="_blank"}

- NOTE: The order of the material varies slightly between
  semesters so the subjects covered in the syllabus may be
  updated as the semester progresses. I will always let you
  know if there are substantial changes.

[**Fall 2026 Absence Explanation
Form**](https://forms.gle/KouC7MXrnA9pj7vb7){target="_blank"}

- If you are absent or have a temporary technical issue with
  POLL EVERYWHERE, use this Google form to explain.

[**Fall 2026 MAS 261 Blackboard
Site**](https://blackboard.syracuse.edu/ultra/courses/_567317_1/outline){target="_blank"}

- Engagement Questions, Assignments, and Tests will be
  submitted on Blackboard.

[**Posit Cloud Signup
Page**](https://posit.cloud/plans/free){target="_blank"}

- Use this link to sign up for a **FREE** Posit Cloud
  account.

[**Read this letter about
AI**](https://www.nytimes.com/2025/06/18/learning/we-need-to-chat-gpt.html?smid=url-share){target="_blank"}

- Using AI in coding and writing is commonplace and helpful
  if used cautiously and ethically.

- This is a great 'letter' from a high school student who
  'broke up' with AI because she felt it was supplanting her
  own voice. It's worth your time to read and consider it.

**Useful Software Links**

- [Posit Cloud](https://posit.cloud/){target="_blank"}

- [R](https://cran.r-project.org/){target="_blank"}

- [RStudio](https://posit.co/products/open-source/rstudio){target="_blank"}

- [Quarto](https://quarto.org/docs/get-started/){target="_blank"}

**OPTIONAL Software Training**

- [LinkedIn
  Learning](https://linkedinlearning.syr.edu/){target="_blank"} -
  Free for Syracuse University Students, Staff, and Faculty

  - [R for Data Science: Analysis and
    Visualization](https://www.linkedin.com/learning/r-for-data-science-analysis-and-visualization?u=76816226){target="_blank"}

    - This training is optional and provides a good
      introduction to R and RStudio which are used in this
      course.

  - [Python Essential
    Training](https://www.linkedin.com/learning/python-essential-training-18764650?u=76816226){target="_blank"}

    - Python is not used in this course but it is helpful
      for analytics and data science.

<br>

- [DataCamp](https://www.datacamp.com/){target="_blank"}

  - An excellent (but not free) online learning center for
    many software packages. Costs can be subsidized (see
    below).

<br>

- [Whitman WIRE
  Initiative](https://whitman.syracuse.edu/experiential-learning/wire-initiative){target="_blank"}

  - The WIRE initiative will subsidize costs of online
    software training that results in certification.

  - Contact their office to arrange a meeting and develop a
    plan.

<br>

- [SU Open Source Program Office
  (OSPO)](https://opensource.syracuse.edu/){target="_blank"}

  - OSPO offers workshops in open source software such as
    R/RStudio, Python, Github and facilitates communication
    between open source software users on campus.

# Lectures 1-12

## Row {height="45%"}

[**MAS 261 Student Files for Lectures 1-12 - R Project on Posit Cloud**](https://posit.cloud/content/10735606){target="_blank"} - Click link to open file in your Posit Cloud and save it. Once file is saved to your account, you should not click this link again.

**NOTE:**  These videos were created in the FALL of 2025 when the first section of the course was lectures 1-8 instead of lectures 1-12.

- [**VIDEO - Creating a FREE Posit Cloud Account**](https://video.syr.edu/media/t/1_a0qm1usm/352723432){target="_blank"}
  
- [**VIDEO - Intro to Accessing Student Lecture R files in Posit Cloud**](https://video.syr.edu/media/t/1_btulm9h8/352723432){target="_blank"}

- [**VIDEO - OPTIONAL Useful Modifications to the Global Options in RStudio**](https://video.syr.edu/media/t/1_bx1cmhms/352723432){target="_blank"}

- [**VIDEO - Navigating an R File and Running Code in an R .qmd File in Posit Cloud**](https://video.syr.edu/media/t/1_0pj3i35n/352723432){target="_blank"}

- [**VIDEO - OPTIONAL - Rendering and Printing an R .qmd File in Posit Cloud**](https://video.syr.edu/media/t/1_b1kccc7t/352723432){target="_blank"}

<br>

**Note:** To make formatted tables in your .qmd more visible
  in the file output add the following text after
  **`kable()`**

  - **`|> kable_styling(full_width = F)`**

## Row {height="55%"}

[***Poll
Everywhere***](https://pollev.com/penelopepoolereisenbies685){target="_blank"
style="color: lightgreen"} - My User Name:
**penelopepoolereisenbies685**

**This website was last updated on `r today_fmt`.**

<br>

**Describing Data**

<br>

- **Lecture 1 - Course Introductions and Types of Data**

  - [**Slides**](https://019f90f0-8209-be03-a14e-6828de956e8b.share.connect.posit.cloud/){target="_blank"}

  - [**Notes**](https://019f90e4-1413-83dc-5528-35b3666f1bca.share.connect.posit.cloud/){target="_blank"}

  - [**PDF**](https://drive.google.com/file/d/1mbjjq4OMmBb81ELnYVK73EVp0nIowyX8/view?usp=sharing){target="_blank"}

  - [**Posit Cloud**](https://posit.cloud/){target="_blank"}

    - Your saved file with your work will be in your Posit Cloud account.
      account.

# Lectures 16-26

## Row {height="35%"}

**The link for the R Project with MAS 261 Student Files for
Lectures 16-26 will be posted before Lecture 16.**

<br>

- Note: To make formatted tables in your .qmd more visible
  in the file output add the following text after
  **`kable()`**

  - **`|> kable_styling(full_width = F)`**

## Row {height="65%"}

[***Poll
Everywhere***](https://pollev.com/penelopepoolereisenbies685){target="_blank"
style="color: lightgreen"} - My User Name:
**penelopepoolereisenbies685**

**This website was last updated on `r today_fmt`.**

# Homework Assignments

## Row

[**Homework Assignments 2-5 - R Project on Posit Cloud**](https://posit.cloud/content/8556337){target="_blank"}

- This R project is now available.

- There is an R Quarto (`.qmd`) file for each Homework
  Assignment that will help you answer the questions in the
  associated Blackboard assignment.
  
- Demo videos will be posted to assist you.

 - **NOTE:** The videos were created in 2025 when the first set of HW files was for assignments 2-4, not 2-5.

<br>

## Row

**HW 2** - Answer questions on Blackboard.

- [**Posit Cloud**](https://posit.cloud/){target="_blank"}

  - Your saved file with your work will be in your account once you click the link for assignments 2-5 at the top of this page..

- [**HW 2 Demo Video Playlist**](https://video.syr.edu/playlist/dedicated/352723432/1_3dpa5815/){target="_blank"}

  - This video playlist was created in the fall of 2025.  If notice any issues, other than the incorrect '2023' label at the top of the assignment, please let me know.

# Installing R and Rstudio

## Row

**Notes**

- Students in MAS 261 are NOT required to install R and
  RStudio on their laptop but they are welcome to.

- R and RStudio are two separate software components.
  Download and install R and then download and install
  RStudio.

- When updating your version of R or RStudio, uninstall the
  previous version first.

## Row

**Downloading and Installing R**

- R can be downloaded from the [**Comprehensive R Archive
  Network**](https://cran.r-project.org/){target="_blank"}

- [**Video Demo for downloading and installing R on a
  Windows
  OS**](https://video.syr.edu/media/t/1_m775pb4g/352723432){target="_blank"}

- Mac installation is similar.

- If you do not need to uninstall an out-of-date version of
  R, skip to 0:53.

- The current version of R is **`4.6.1`**.

<br>

**Downloading and Installing RStudio**

- RStudio can be downloaded from
  [**Posit**](https://posit.co/download/rstudio-desktop/){target="_blank"},
  the parent company of RStudio.

- [**Video Demo for downloading and installing RStudio on a
  Windows
  OS**](https://video.syr.edu/media/t/1_my39umi3/352723432){target="_blank"}

- Mac installation is similar.

- If you do not need to uninstall an out-of-date version of
  RStudio, skip to 0:56.

- The version shown in the video is from August of 2024 and
  is out-of-date.

- The current version of RStudio is **`2026.8.1.195`**.
